Emissions and Perceptions of Climate Change

### 2022-12-14 Code/FINAL_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Cleaning up the Markdown

```{r}
emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
#format tells R the dates are just a four digit year, a two digit year is %y, origin informs R where it starts.
# must do the - in the %d- because the dates are entered with a -. 
```

@@

Lets see the emissions of Denmark and the United States from 1990 to 2019.
```{r}
ggplot(emissions, aes(year, emissions, colour= country, group=country)) +  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.

@@

```{r}
t.test(emissions~country, data=emissions, paired=TRUE)
#like running a t test, but tell r the data is paired.
```
The p-value is very less than 2.2*10^-16, much less than 0.05, so we can reject our Null hypothesis. Denmark does have significantly less GHG emissions per capita than the United States.

@@

Lets look at the rate of Denmark and the United States.
```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
```
Figure 2. The rate of per capita greenhouse gas emissions for Denmark and the U.S. from 1990 to 2019.

@@

```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
  geom_smooth() + theme_bw() #colour colors by country, and group tells R to 
#group the points by country so geom_line will draw lines connecting the 
#data by country.

@@

### Running the Statistical Test
Lets run a paired t-test to see if this is significant.
```{r}
t.test(rate~country, data=emissions_rate, paired=TRUE)
#complete.cases means just use the ones without NAs
```

Looking at the entire range from 1990-2019, the p value is 0.7668, so we must accept the Null hypothesiss. There is no significant difference in rate between Denmark and the U.S. Lets look at 2010-2019 given the differences in the funcitons listed above.

@@

#Resetting the levels

beliefs$response<-factor(beliefs$response)
levels(beliefs$response)
```

@@

```{r}
# Making the graph
ggplot(beliefs, aes(response,Number, fill=Country))+
  geom_bar(stat = 'identity', position = 'dodge') + 
 labs(title="Cause of Climate Change", x="Response to Survey Question", y="Percent") + # titling and renaming x and y axis labels
  #theme( axis.text.x = element_text(angle = 45, size = 13, vjust = 0.5))+ this won't work :(
  coord_flip()+
  theme_bw()

@@

```{r}
beliefs_diverging <- beliefs %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Caused mostly by human activities"), Number, -Number)) 
#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
```

```{r}
beliefs_diverging_Order<-beliefs_diverging %>% mutate(response=fct_relevel(response, "Caused about equally by human activities and natural changes", "Other", "Caused mostly by natural changes in the environment", "None of the above because climate change isn’t happening", "Caused mostly by human activities"), response=fct_rev(response))

beliefs_diverging_Order
```

```{r}

#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
beliefs_diverging_Order<- beliefs_diverging_Order %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))

@@

  #reorganizes the labels and color them
  scale_fill_manual(breaks=c("None of the above because climate change isn’t happening","Caused mostly by natural changes in the environment", "Other", "Caused about equally by human activities and natural changes","Caused mostly by human activities"), values=c(
    "None of the above because climate change isn’t happening"="darkorange4",
    "Caused mostly by natural changes in the environment" = "darkorange3",
    "Other" = "darkorange", 
    "Caused about equally by human activities and natural changes"= "goldenrod1",
    "Caused mostly by human activities" = "deepskyblue"
  ))+

@@

```{r, echo=FALSE}
respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
beliefs$num_people<- respondants #put that vector into here

@@

```{r}

beliefs_numPeople$grouped_answer<- NA
good<- which( beliefs_numPeople$response== "Caused mostly by human activities")
bad<-which(beliefs_numPeople$response=="Caused mostly by natural changes in the environment" |beliefs_numPeople$response=="Caused about equally by human activities and natural changes"
           |beliefs_numPeople$response=="Other"
           |beliefs_numPeople$response=="None of the above because climate change isn’t happening")
  beliefs_numPeople$grouped_answer[good]<- "good"
  beliefs_numPeople$grouped_answer[bad]<-"bad"

@@

```{r}
beliefs.mat<-xtabs(grouped_number~Country + grouped_answer, data=beliefs_numPeople)
beliefs.mat
chisq.test(beliefs.mat)
```
The p values is 3.829*10^-11, which is much less than 95%, so I can reject the Null hypothesis. Significantly more Danes think climate change is caused by humans verses other sources than Americans.

@@

```{r}
ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
```

@@

Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
```{r}
worry_diverging <- worry %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very worried", "Somewhat worried"), Number, -Number)) 
#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
```

```{r}
worry_diverging<-worry_diverging %>% mutate(response=fct_relevel(response, "Not at all worried","Not very worried", "Somewhat worried", "Very worried"), response=fct_rev(response))

@@

```{r}

@@

```{r}

worry_numPeople<- worry_numPeople %>% filter(response!="Refused")
worry_numPeople$grouped_answer<- NA
not<- which(worry_numPeople$response== "Not at all worried"|worry_numPeople$response== "Not very worried")
worried<- which( worry_numPeople$response== "Very worried" |worry_numPeople$response== "Somewhat worried")

@@

```{r}
worry.mat<-xtabs(grouped_number~Country+grouped_answer, data=worry_numPeople)
worry.mat
chisq.test(worry.mat)
```
the p-value is 0.01258, which is less than 0.05. So, we can reject the Null hypothesis. Danes are more worried about climate change than Americans.

@@

```{r}
ggplot(threat, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  labs(title="Perceieved Threat of Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
```

@@

Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
```{r}
threat_diverging <- threat %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very serious threat", "Somewhat serious threat"), Number, -Number))

@@

Reorganize the bars.
```{r}
threat_diverging<-threat_diverging %>% mutate(response=fct_relevel(response, "Don't know", "Not a threat at all", "Somewhat serious threat", "Very serious threat"), response=fct_rev(response))

@@

```{r}
#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
threat_diverging<- threat_diverging %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))

@@

```{r}

@@

```{r}
# Making the graph
ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  labs(title="Fossil Fuel Use", x="Response to Survey Question", y="Percent") + # renaming x and y axis labels
  theme_bw() +
  coord_flip() # switching x and y axis to read the suvey questions
```

@@

Again the option "Don't know" is present. This also has the option of same amount as today. According to the IPCC humans must decrease their emissions to stay well below 2 degrees of warming and avoid the most severe impacts of climate change, therefore, keeping it the same will go on the negative side of the diverging bar graph. Not many people chose "Don't know," plus it is difficult to predict, so I believe it is fair to leave it out.
```{r}
fuels_diverging <- fuels %>% filter(response!="Refused" & response!= "Don’t know") %>% mutate(Number=if_else(response %in% c("Much less", "Somewhat less"), Number, -Number))

@@

Reorganize the bars.
```{r}
fuels_diverging<-fuels_diverging %>% mutate(response=fct_relevel(response,"Much more","Somewhat more", "Same amount as today", "Somewhat less", "Much less"), response=fct_rev(response))

```{r}
#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(abs(Number), digits=0), "%"))
#rounded it to 0 this time due to serious overlap in the % labels on the graph

@@

fuels_numPeople<- fuels_numPeople %>% filter(response!="Refused" & response!="Don't know") 
fuels_numPeople$grouped_answer<- NA
more<- which(fuels_numPeople$response== "Much more" | fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Same amount as today" )
less<- which(fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less")

>>>>

```{r}
emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
#format tells R the dates are just a four digit year, a two digit year 
#is %y, origin informs R where it starts.
# must do the - in the %d- because the dates are entered with a -. 
```

@@

Lets see the emissions of Denmark and the United States from 1990 to 2019.
```{r}
ggplot(emissions, aes(year, emissions, colour= country, group=country)) +  
  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, 
#and group tells R to group the points by country so geom_line will draw
#lines connecting the data by country.

@@

```{r, results='hold'}
t.test(emissions~country, data=emissions, paired=TRUE)
#like running a t test, but tell r the data is paired.
```
The p-value is very less than 2.2*10^-16, much less than 0.05, so we can reject our Null hypothesis. Denmark does have significantly less GHG emissions per capita than the United States.

@@

Lets look at the rate of Denmark and the United States.
```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) + 
  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, 
#and group tells R to group the points by country so geom_line will draw 
#lines connecting the data by country.
```
Figure 2. The rate of per capita greenhouse gas emissions for Denmark and the U.S. from 1990 to 2019.

@@

```{r, message=FALSE}
#colour colors by country, and group tells R to 
#group the points by country so geom_line will draw lines connecting the 
#data by country.
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
  geom_smooth() + theme_bw()

@@

### Running the Statistical Test
Lets run a paired t-test to see if this is significant.
```{r}
#complete.cases means just use the ones without NAs
t.test(rate~country, data=emissions_rate, paired=TRUE)
```

Looking at the entire range from 1990-2019, the p value is 0.7668, so we must accept the Null hypothesiss. There is no significant difference in rate between Denmark and the U.S. Lets look at 2010-2019 given the differences in the functions listed above.

@@

#Resetting the levels
beliefs$response<-factor(beliefs$response)
levels(beliefs$response)
```

@@

```{r, warning=FALSE}
# Making the graph
ggplot(beliefs, aes(response,Number, fill=Country))+
  geom_bar(stat = 'identity', position = 'dodge') + 
 labs(title="Cause of Climate Change", x="Response to Survey Question", 
      y="Percent") + # titling and renaming x and y axis labels
  coord_flip()+
  theme_bw()

@@

```{r}
beliefs_diverging <- beliefs %>% filter(response!="Refused") %>% 
  mutate(Number=if_else(response %in% 
                    c("Caused mostly by human activities"), Number, -Number)) 
#all the percentages in the response column that did not 
#answer "caused mostly by human activities"   will have a - put in front. 
```

```{r}
beliefs_diverging_Order<-beliefs_diverging %>% 
  mutate(response=fct_relevel(response, 
  "Caused about equally by human activities and natural changes", 
  "Other", "Caused mostly by natural changes in the environment", 
  "None of the above because climate change isn’t happening", 
  "Caused mostly by human activities"), response=fct_rev(response)) 
```

```{r, warning=FALSE, message=FALSE}

#first line below takes the Number column rounds it up, 
#takes the absolute value of it and adds a % for the labels on the graph.
beliefs_diverging_Order<- beliefs_diverging_Order %>%
  mutate(percent=paste0(round(abs(Number), digits=0), "%"))

@@

  #reorganizes the labels and color them
  scale_fill_manual(breaks=c("None of the above because climate change isn’t happening",
  "Caused mostly by natural changes in the environment", "Other", 
  "Caused about equally by human activities and natural changes",
  "Caused mostly by human activities"), 
  values=c(
    "None of the above because climate change isn’t happening"="darkorange4",
    "Caused mostly by natural changes in the environment" = "darkorange3",
    "Other" = "darkorange", 
    "Caused about equally by human activities and natural changes"= "goldenrod1",
    "Caused mostly by human activities" = "deepskyblue"
  ))+

@@

Given  the diverging bar graph and chi squared test code chunks are virtually the same as above, they are not shown in the pdf version henceforth. 
```{r, echo=FALSE}
respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
beliefs$num_people<- respondants #put that vector into here

@@

beliefs_numPeople$grouped_answer<- NA
good<- which( beliefs_numPeople$response== "Caused mostly by human activities")
bad<-which(beliefs_numPeople$response==
             "Caused mostly by natural changes in the environment"
|beliefs_numPeople$response=="Caused about equally by human activities and natural changes"
           |beliefs_numPeople$response=="Other"
|beliefs_numPeople$response=="None of the above because climate change isn’t happening")
  beliefs_numPeople$grouped_answer[good]<- "good"
  beliefs_numPeople$grouped_answer[bad]<-"bad"

@@

```{r, echo=FALSE}
beliefs.mat<-xtabs(grouped_number~Country + grouped_answer, data=beliefs_numPeople)
beliefs.mat
chisq.test(beliefs.mat)
```
The p values is 3.829*10^-11, which is much less than 95%, so I can reject the Null hypothesis. Significantly more Danes think climate change is caused by humans verses other sources than Americans.

@@

```{r, echo=FALSE}
ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity',
                                                          position = 'dodge') + 
  labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", 
       y="Percent") + 
  theme_bw() +
  coord_flip()
```

@@

Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
```{r, echo=FALSE}
worry_diverging <- worry %>% filter(response!="Refused") %>% 
  mutate(Number=if_else(response %in% c("Very worried", "Somewhat worried"), 
                        Number, -Number))

@@

```{r, echo=FALSE}
worry_diverging<-worry_diverging %>% mutate(response=fct_relevel(response,
  "Not at all worried","Not very worried", "Somewhat worried", "Very worried"), 
  response=fct_rev(response))

@@

worry_numPeople<- worry_numPeople %>% filter(response!="Refused")
worry_numPeople$grouped_answer<- NA
not<- which(worry_numPeople$response== "Not at all worried"
            |worry_numPeople$response== "Not very worried")
worried<- which( worry_numPeople$response== "Very worried" 
                 |worry_numPeople$response== "Somewhat worried")

@@

```{r, echo=FALSE}
worry.mat<-xtabs(grouped_number~Country+grouped_answer, data=worry_numPeople)
worry.mat
chisq.test(worry.mat)
```
the p-value is 0.01258, which is less than 0.05. So, we can reject the Null hypothesis. Danes are more worried about climate change than Americans.

@@

```{r, echo=FALSE}
ggplot(threat, aes(response,Number, fill=Country))+geom_bar(stat = 'identity',
                                                        position = 'dodge') + 
  labs(title="Perceieved Threat of Climate Change", 
       x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
```

@@

Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
```{r, echo=FALSE}
threat_diverging <- threat %>% filter(response!="Refused") %>% 
  mutate(Number=if_else(response %in% c("Very serious threat", 
                                "Somewhat serious threat"), Number, -Number))

@@

```{r, echo=FALSE}
threat_diverging<-threat_diverging %>% mutate(response=fct_relevel(response,
"Don't know", "Not a threat at all", "Somewhat serious threat", "Very serious threat"), response=fct_rev(response))

@@

```{r, echo=FALSE}
#first line below takes the Number column rounds it up, takes the absolute 
#value of it and adds a % for the labels on the graph.
threat_diverging<- threat_diverging %>% mutate(percent=paste0(round(abs(Number), 
                                                              digits=1), "%"))

@@

@@

```{r, warning=FALSE, echo=FALSE}
# Making the graph
ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity',
                                                           position = 'dodge') + 
  labs(title="Fossil Fuel Use", x="Response to Survey Question", y="Percent") + 
  # renaming x and y axis labels
  theme_bw() +
  coord_flip() # switching x and y axis to read the suvey questions
```

@@

Again the option "Don't know" is present. This also has the option of same amount as today. According to the IPCC humans must decrease their emissions to stay well below 2 degrees of warming and avoid the most severe impacts of climate change, therefore, keeping it the same will go on the negative side of the diverging bar graph. Not many people chose "Don't know," plus it is difficult to predict, so I believe it is fair to leave it out.
```{r, echo=FALSE}
fuels_diverging <- fuels %>% filter(response!="Refused" 
                                    & response!= "Don’t know")%>% 
  mutate(Number=if_else(response %in% c("Much less", "Somewhat less"), 
                          Number, -Number)) 
```

```{r, echo=FALSE}
fuels_diverging<-fuels_diverging %>% mutate(response=fct_relevel(response,
"Much more","Somewhat more", "Same amount as today", "Somewhat less", 
"Much less"), response=fct_rev(response))

@@

```{r,echo=FALSE}
#first line below takes the Number column rounds it up, takes the absolute 
#value of it and adds a % for the labels on the graph.
fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(abs(Number),
                                                            digits=0), "%"))
#rounded it to 0 this time due to serious overlap in the % labels on the graph

@@

fuels_numPeople<- fuels_numPeople %>% filter(response!="Refused" & 
                                               response!="Don't know") 
fuels_numPeople$grouped_answer<- NA
more<- which(fuels_numPeople$response== "Much more" |
  fuels_numPeople$response== "Somewhat more" | 
    fuels_numPeople$response== "Same amount as today" )
less<- which(fuels_numPeople$response== "Much less" |
               fuels_numPeople$response== "Somewhat less")
